---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 
```{r}
library(tidyverse)
```


Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
deaths <- av %>% 
  pivot_longer(
    starts_with("Death"),
    names_to = "Time",
    values_to = "Died"
  ) %>% mutate(
    Time = parse_number(Time)
  ) %>% pivot_longer(
    starts_with("Return"),
    names_to = 'Time_Returned',
    values_to = 'Returned'
  ) %>% mutate(
    Time_Returned = parse_number(Time_Returned)
  )
```

```{r}
average_deaths <- deaths %>%
  group_by(Name.Alias) %>% 
  summarise(death_count = sum(Died == "YES", na.rm = TRUE)) %>%
  summarise(avg_deaths = mean(death_count, na.rm = TRUE))

print(average_deaths)
```


## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

Akshat - "There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife"

Ryan - "Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.

Jacob - "I counted 89 total deaths and on 57 occasions the individual made a comeback."

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}
#Akshat's Code
avenger_deaths <- deaths %>%
  filter(Died == "YES") %>%
  group_by(Name.Alias) %>%
  mutate(death_number = row_number()) %>%
  ungroup()

return_rates <- avenger_deaths %>%
  filter(death_number == 1) %>%
  group_by(death_number) %>%
  summarise(return_rate = mean(Returned == "YES", na.rm = TRUE))

return_rates
```


Ryan's Code

``` {r}
av %>%
  filter(Death1 == 'YES') %>%
  summarise(total_died = n())
```

### Jacob's Code
```{r}
deaths_count <- av %>% 
  pivot_longer(
    starts_with("Death"),
    names_to = "Time",
    values_to = "Died"
  ) %>% mutate(
    Time = parse_number(Time)
  )
returns <- av %>% 
    pivot_longer(
    starts_with("Return"),
    names_to = 'Time_Returned',
    values_to = 'Returned'
  ) %>% mutate(
    Time_Returned = parse_number(Time_Returned)
  )

deaths_count %>%
  filter(Died == "YES") %>%
  summarise(total_died = n())

returns %>%
  filter(Returned == "YES") %>%
  summarise(total_returned = n())
```

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.


Akshat - The statement that "There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife" is approximately correct as there is a 70% chance of return after the first death and 2/3 is approximtely 66.67%.

Ryan's Result: My findings match the statement on the FiveThirtyEight website, as I also found that 69 avengers died at least one time after they joined the team.

Jacob - My findings match the statement from the article that there's 89 total deaths and 57 instances of an individual returning.

Upload your changes to the repository. Discuss and refine answers as a team.
